---
title: "Project Proposal"
author: "Chen Si (chensi3), Yuchen Cao (yuchenc5)"
date: "2023-03-31"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaps)
library(ggplot2)
library(dplyr)
```

### Read in Dataset
```{r}
fifa<-read.csv("../dataset/players_20_edited.csv",header = T,stringsAsFactors = T)

colnames(fifa)

set.seed(1) # for part (d), you change the random seed.
train.id=sample(1:nrow(fifa),trunc(0.9*nrow(fifa)))
tr.col=fifa[train.id,] #training data
tst.col=fifa[-train.id,] #test data

est.id=sample(1:nrow(tr.col),trunc(0.9*nrow(tr.col)))
est.col=tr.col[est.id,] #estimation data
val.col=tr.col[-est.id,] #validation data
```
  
### Linear Regression

Perform the best subset selection on the estimation data. Report the model choice based on Cp criteria.

```{r}
RSSleaps=regsubsets(overall ~ age + height_cm + weight_kg +
                    shooting + shooting + passing + dribbling +
                    defending + physic + attacking_crossing +
                    attacking_finishing + attacking_heading_accuracy +
                    attacking_short_passing + attacking_volleys +
                    skill_dribbling + skill_curve + skill_fk_accuracy +
                    skill_long_passing + skill_ball_control +
                    movement_acceleration + movement_sprint_speed +
                    movement_agility + movement_reactions +
                    movement_balance + power_shot_power +
                    power_jumping + power_stamina + power_strength +
                    power_long_shots + mentality_aggression +
                    mentality_interceptions + mentality_positioning +
                    mentality_vision + mentality_penalties +
                    mentality_composure + defending_marking +
                    defending_standing_tackle + defending_sliding_tackle,data=est.col,nvmax=37)

sumleaps=summary(RSSleaps,matrix=T)
modelsize=apply(sumleaps$which,1,sum)

plot(modelsize,sumleaps$cp)

which.min(sumleaps$cp)

#model choice:
coef(RSSleaps,24)
```

**Report**
The model we choose is overall ~ age + weight_kg + shooting + passing + defending + physic + attacking_heading_accuracy + attacking_short_passing + skill_fk_accuracy + skill_long_passing + skill_ball_control + movement_acceleration + movement_sprint_speed + movement_agility + movement_reactions + movement_balance + power_long_shots + mentality_aggression + mentality_interceptions + mentality_positioning + mentality_vision + mentality_composure + defending_marking + defending_sliding_tackle.

Perform the backward selection on the estimation data. Report the model choice based on BIC criteria. 

```{r}
lm.fit=lm(overall ~ age + height_cm + weight_kg + 
                    shooting + shooting + passing + dribbling + 
                    defending + physic + attacking_crossing +
                    attacking_finishing + attacking_heading_accuracy +
                    attacking_short_passing + attacking_volleys +
                    skill_dribbling + skill_curve + skill_fk_accuracy +
                    skill_long_passing + skill_ball_control +
                    movement_acceleration + movement_sprint_speed +
                    movement_agility + movement_reactions +
                    movement_balance + power_shot_power + 
                    power_jumping + power_stamina + power_strength +
                    power_long_shots + mentality_aggression +
                    mentality_interceptions + mentality_positioning +
                    mentality_vision + mentality_penalties +
                    mentality_composure + defending_marking +
                    defending_standing_tackle + defending_sliding_tackle,data=est.col)
n=nrow(est.col)
step(lm.fit,direction="backward",k=log(n))
```

**Report**
The model we choose is overall ~ age + weight_kg + passing + defending + attacking_finishing + attacking_heading_accuracy + attacking_short_passing + skill_long_passing + skill_ball_control + movement_acceleration + movement_sprint_speed + movement_reactions + movement_balance + power_shot_power + power_stamina + power_strength + power_long_shots + mentality_interceptions + mentality_positioning + mentality_vision + mentality_composure + defending_sliding_tackle.

Which model gives lowest validation error?

```{r}
lm.fit.a <- lm(overall ~ age + weight_kg + shooting + passing + defending + physic + attacking_heading_accuracy + attacking_short_passing + skill_fk_accuracy + skill_long_passing + skill_ball_control + movement_acceleration + movement_sprint_speed + movement_agility + movement_reactions + movement_balance + power_long_shots + mentality_aggression + mentality_interceptions + mentality_positioning + mentality_vision + mentality_composure + defending_marking + defending_sliding_tackle,data = est.col)

lm.fit.b <- lm(overall ~ age + weight_kg + passing + defending + attacking_finishing + attacking_heading_accuracy + attacking_short_passing + skill_long_passing + skill_ball_control + movement_acceleration + movement_sprint_speed + movement_reactions + movement_balance + power_shot_power + power_stamina + power_strength + power_long_shots + mentality_interceptions + mentality_positioning + mentality_vision + mentality_composure + defending_sliding_tackle,data = est.col)

val.pred.a=predict(lm.fit.a,val.col) 
val.pred.b=predict(lm.fit.b,val.col)
mean((val.pred.a-val.col$overall)^2)
mean((val.pred.b-val.col$overall)^2)
```

**Report**
We choose the second model.

```{r}
lm.fit=lm(overall ~ age + height_cm + weight_kg + 
                    shooting + shooting + passing + dribbling + 
                    defending + physic + attacking_crossing +
                    attacking_finishing + attacking_heading_accuracy +
                    attacking_short_passing + attacking_volleys +
                    skill_dribbling + skill_curve + skill_fk_accuracy +
                    skill_long_passing + skill_ball_control +
                    movement_acceleration + movement_sprint_speed +
                    movement_agility + movement_reactions +
                    movement_balance + power_shot_power + 
                    power_jumping + power_stamina + power_strength +
                    power_long_shots + mentality_aggression +
                    mentality_interceptions + mentality_positioning +
                    mentality_vision + mentality_penalties +
                    mentality_composure + defending_marking +
                    defending_standing_tackle + defending_sliding_tackle,data=fifa)
ggplot(fifa, aes(x = overall, y = predict(lm.fit))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Actual Overall", y = "Predicted Overall") +
  ggtitle("Linear Regression of FIFA2020 Data")

#training MSE
final.model=lm(overall ~ age + height_cm + weight_kg + 
                    shooting + shooting + passing + dribbling + 
                    defending + physic + attacking_crossing +
                    attacking_finishing + attacking_heading_accuracy +
                    attacking_short_passing + attacking_volleys +
                    skill_dribbling + skill_curve + skill_fk_accuracy +
                    skill_long_passing + skill_ball_control +
                    movement_acceleration + movement_sprint_speed +
                    movement_agility + movement_reactions +
                    movement_balance + power_shot_power + 
                    power_jumping + power_stamina + power_strength +
                    power_long_shots + mentality_aggression +
                    mentality_interceptions + mentality_positioning +
                    mentality_vision + mentality_penalties +
                    mentality_composure + defending_marking +
                    defending_standing_tackle + defending_sliding_tackle,data=tr.col)
train.pred=predict(final.model,tr.col)
mean((train.pred-tr.col$overall)^2)

#test MSE
test.pred=predict(final.model,tst.col)
mean((test.pred-tst.col$overall)^2)
```

### Logistic Regression

First, we will bin the player values and name the new column as Value Brackets with the labels: 0–10M, 10–20M, 20–30M, 30–40M, 40–50M, 50–60M, 60–70M, 70–80M, 80–90M, 90–100M, 100M+.

```{r}
# Create value brackets
value_breaks <- c(0, 10000000, 20000000, 30000000, 40000000, 50000000, 60000000, 70000000, 80000000, 90000000, 100000000, Inf)
value_labels <- c("0–10M", "10–20M", "20–30M", "30–40M", "40–50M","50–60M", "60–70M", "70–80M", "80–90M","90–100M","100M+")

`Value Brackets` <- cut(x=fifa$value_eur, breaks=value_breaks, 
                      labels=value_labels, 
                      include.lowest = TRUE)

fifa <- mutate(fifa, `Value Brackets`)
```

We see that the high valuations are dominated by players of overall rating 85+ and age between 23 to 33 years.

```{r}
g_age_overall <- ggplot(fifa, aes(age, overall))
g_age_overall + 
  geom_point(aes(color = `Value Brackets`)) + 
  geom_smooth(color = "darkblue") + 
  ggtitle("Distribution between Age and Overall Rating of players based on Value bracket") + 
  theme_minimal()
```

Now, we will perform logistic regression analysis to find the relationship between Overall and Age.

```{r}
fifa$good_player <- ifelse(fifa$overall >= 75, 1, 0)
model <- glm(good_player ~ age, data=fifa, family="binomial")

summary(model)
```

Plot the logistic regression graph.

```{r}
ggplot(fifa, aes(x=age, y=good_player)) +
  geom_point() +
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE, color="red") +
  ggtitle("Logistic Regression Curve for Age vs Good Player")
```